---
title: "Untitled"
output:
  html_document: default
  pdf_document: default
date: "2024-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## World Population Analysis

Comprehensive World Population EDA

```{r list:files}
list.files(path = "archive")
```

```{r libraries}
library(skimr)
library(janitor)
library(ggplot2)
library(reshape2)
library(ggpie)
library(plotrix)
library(gridExtra)
#library(tidyverse)
#install.packages("dplyr")
library(dplyr)
```

## Importing the dataset

```{r dataset, echo=FALSE}
world_pop <- read.csv("./archive/world_population.csv")
colnames(world_pop)
```

```{r}
#renaming the column names for the dataset
colnames(world_pop)[6] <- "population_2022"
colnames(world_pop)[7] <- "population_2020"
colnames(world_pop)[8] <- "population_2015"
colnames(world_pop)[9] <- "population_2010"
colnames(world_pop)[10] <- "population_2000"
colnames(world_pop)[11] <- "population_1990"
colnames(world_pop)[12] <- "population_1980"
colnames(world_pop)[13] <- "population_1970"
colnames(world_pop)[14] <- "square_area"
colnames(world_pop)[15] <- "pop_density"
colnames(world_pop)[16] <- "growth_rate"
colnames(world_pop)[17] <- "world_pop_percent"
```

```{r}
#checking the column names
#checking the dataset's summary
colnames(world_pop)
sum(is.na(world_pop))
str(world_pop)
```

There are no missing values in this data set and the column names have been changed as well.

### Visualizing The Number of Countries Per Continent

```{r}
#number of countries per continent
country_count <- world_pop %>% 
  group_by(Continent) %>%
  summarise(Count = n())
countries_per_continent <-ggplot(country_count, aes(x=Continent, y=Count, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= Count), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Countries Count', title = 'Countries per Continent')+
  theme(axis.text.x = element_text(angle = 45))
countries_per_continent
```

-   Africa has the most number of countries and South America has the least amount of countries

Total Population by Year by Continent¶

Year 2022

```{r}
pop_2022 <- world_pop %>%
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2022)) %>%
  arrange(desc(total_pop))

pop2022 <- ggplot(pop_2022, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2022)')+
  theme(axis.text.x = element_text(angle = 45))
pop2022
```

-   In the year 2022, Asia has the highest total population of all the continents followed by Africa.

Year 2020

```{r}
# 2020
pop_2020 <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2020)) %>%
  arrange(desc(total_pop))

pop2020 <- ggplot(pop_2020, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2020)')+
  theme(axis.text.x = element_text(angle = 45))
pop2020
```

Similar to the year 2022, Asia has the highest total population of all the continents followed by Africa.

Year 2015

```{r}
# 2015
pop_2015 <- world_pop %>%
  group_by(Continent) %>%
  summarise(total_pop = sum(population_2015)) %>%
  arrange(desc(total_pop))

pop2015 <- ggplot(pop_2015, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2015)')+
  theme(axis.text.x = element_text(angle = 45))
pop2015

```

Year 2010

```{r}
# 2010
pop_2010 <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2010)) %>%
  arrange(desc(total_pop))

pop2010 <- ggplot(pop_2010, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2010)')+
  theme(axis.text.x = element_text(angle = 45))
pop2010
```

### Year 2000

```{r}
# 2000
pop_2000 <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2000)) %>%
  arrange(desc(total_pop))

pop2000 <- ggplot(pop_2000, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2000)')+
  theme(axis.text.x = element_text(angle = 45))
print(pop_2000)
```

-   For the year 2000, the population for Africa exceeded that of Europe as it continues in its growth rate keeps rising compared to that of Europe, Asia still has the highest population.

### Total Population Trend by Continent

Perfoming data Manipulation \`{r}

```{r}
# Total population trend by continent
# Performing some data manipulation to create a new table to place the continents as columns and the total population as rows
pop_combined <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop_2022 = sum(population_2022), total_pop_2020 = sum(population_2020),
            total_pop_2015 = sum(population_2015), total_pop_2010 = sum(population_2010),
            total_pop_2000 = sum(population_2000), total_pop_1990 = sum(population_1990),
            total_pop_1980 = sum(population_1980), total_pop_1970 = sum(population_1970)) 

pop_combined <- t(select(pop_combined, "total_pop_2022", "total_pop_2020", "total_pop_2015", "total_pop_2010",
       "total_pop_2000", "total_pop_1990", "total_pop_1980", "total_pop_1970"))
```

\`

```{r}
colnames(pop_combined) <- c("Africa", "Asia", "Europe", "North_America", "Oceania", "South_America")
pop_coombined_new <- cbind(Years = rownames(pop_combined), pop_combined)
rownames(pop_coombined_new)<- 1:nrow(pop_coombined_new)
colnames(pop_coombined_new) <- c("Years","Africa", "Asia", "Europe", "North_America", "Oceania", "South_America")
pop_coombined_new <- as.data.frame(pop_coombined_new)
```

```{r}
# Total population trend by continent
# Africa
Africa <- ggplot(pop_coombined_new, aes(x=Years, y=reorder(Years,Africa), fill = Years)) + geom_col(position = "dodge") +
  geom_text(aes(label= Africa), vjust=0, size=3) +
  labs(x= 'Years', y = 'Africa', title = 'Total Population Trend for Africa')+
  theme(axis.text.x = element_text(angle = 45))

# Asia
Asia <-ggplot(pop_coombined_new, aes(x=Years, y=Asia, fill = Years)) + geom_col(position = "dodge") +
  geom_text(aes(label= Asia), vjust=0, size=3) +
  labs(x= 'Years', y = 'Asia', title = 'Total Population Trend for Asia')+
  theme(axis.text.x = element_text(angle = 45))

# Europe
Europe <-ggplot(pop_coombined_new, aes(x=Years, y=Europe, fill = Years)) + geom_col(position = "dodge") +
  geom_text(aes(label= Europe), vjust=0, size=3) +
  labs(x= 'Years', y = 'Europe', title = 'Total Population Trend for Europe')+
  theme(axis.text.x = element_text(angle = 45))

# North America
North_America <-ggplot(pop_coombined_new, aes(x=Years, y=North_America, fill = Years)) + geom_col(position = "dodge") +
  geom_text(aes(label= North_America), vjust=0, size=3) +
  labs(x= 'Years', y = 'North America', title = 'Total Population Trend for North America')+
  theme(axis.text.x = element_text(angle = 45))

# Oceania
Oceania <- ggplot(pop_coombined_new, aes(x=Years, y=Oceania, fill = Years)) + geom_col(position = "dodge") +
  geom_text(aes(label= Oceania), vjust=0, size=3) +
  labs(x= 'Years', y = 'Oceania', title = 'Total Population Trend for Oceania')+
  theme(axis.text.x = element_text(angle = 45))

# South America
South_America <- ggplot(pop_coombined_new, aes(x=Years, y=South_America, fill = Years)) + geom_col(position = "dodge") +
  geom_text(aes(label= South_America), vjust=0, size=3) +
  labs(x= 'Years', y = 'South America', title = 'Total Population Trend for South America')+
  theme(axis.text.x = element_text(angle = 45))
```

-   All the continents over the years from 1970 through 2022 saw different levels of increases in their total population except Europe.

-   Europe's population for the year 2022 saw a drop from the year 2020.

-   Research to find out causes would be interesting, maybe more death rates than birth rates, covid, and so on.

    ```{}
    ```

    ```{}
    ```
