---
title: "Untitled"
output:
  html_document: default
  pdf_document: default
date: "2024-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## World Population Analysis

Comprehensive World Population EDA

```{r list:files}
list.files(path = "archive")
```

```{r libraries}
library(skimr)
library(janitor)
library(ggplot2)
library(reshape2)
library(ggpie)
library(plotrix)
library(gridExtra)
#library(tidyverse)
#install.packages("dplyr")
library(dplyr)
```

## Importing the dataset

```{r dataset, echo=FALSE}
world_pop <- read.csv("./archive/world_population.csv")
colnames(world_pop)
```

```{r}
#renaming the column names for the dataset
colnames(world_pop)[6] <- "population_2022"
colnames(world_pop)[7] <- "population_2020"
colnames(world_pop)[8] <- "population_2015"
colnames(world_pop)[9] <- "population_2010"
colnames(world_pop)[10] <- "population_2000"
colnames(world_pop)[11] <- "population_1990"
colnames(world_pop)[12] <- "population_1980"
colnames(world_pop)[13] <- "population_1970"
colnames(world_pop)[14] <- "square_area"
colnames(world_pop)[15] <- "pop_density"
colnames(world_pop)[16] <- "growth_rate"
colnames(world_pop)[17] <- "world_pop_percent"
```

```{r}
#checking the column names
#checking the dataset's summary
colnames(world_pop)
sum(is.na(world_pop))
str(world_pop)
```

There are no missing values in this data set and the column names have been changed as well.

## Visualizing The Number of Countries Per Continent

```{r}
#number of countries per continent
country_count <- world_pop %>% 
  group_by(Continent) %>%
  summarise(Count = n())
countries_per_continent <-ggplot(country_count, aes(x=Continent, y=Count, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= Count), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Countries Count', title = 'Countries per Continent')+
  theme(axis.text.x = element_text(angle = 45))
countries_per_continent
```

-   Africa has the most number of countries and South America has the least amount of countries

Year 2022 \# Total Population by Year by Continent¶

```{r}
pop_2022 <- world_pop %>%
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2022)) %>%
  arrange(desc(total_pop))

pop2022 <- ggplot(pop_2022, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2022)')+
  theme(axis.text.x = element_text(angle = 45))
pop2022
```

-   In the year 2022, Asia has the highest total population of all the continents followed by Africa.

Year 2020

```{r}
# 2020
pop_2020 <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2020)) %>%
  arrange(desc(total_pop))

pop2020 <- ggplot(pop_2020, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2020)')+
  theme(axis.text.x = element_text(angle = 45))
pop2020
```

Similar to the year 2022, Asia has the highest total population of all the continents followed by Africa.

Year 2015

```{r}
# 2015
pop_2015 <- world_pop %>%
  group_by(Continent) %>%
  summarise(total_pop = sum(population_2015)) %>%
  arrange(desc(total_pop))

pop2015 <- ggplot(pop_2015, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2015)')+
  theme(axis.text.x = element_text(angle = 45))
pop2015

```

Year 2010

```{r}
# 2010
pop_2010 <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2010)) %>%
  arrange(desc(total_pop))

pop2010 <- ggplot(pop_2010, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2010)')+
  theme(axis.text.x = element_text(angle = 45))
pop2010
```

Year 2000

```{r}
# 2000
pop_2000 <- world_pop %>% 
  group_by(Continent) %>% 
  summarise(total_pop = sum(population_2000)) %>%
  arrange(desc(total_pop))

pop2000 <- ggplot(pop_2000, aes(x=Continent, y=total_pop, fill = Continent)) + geom_col(position = "dodge") +
  geom_text(aes(label= total_pop), vjust=0, size=3) +
  labs(x= 'Continent', y = 'Total Population', title = 'Total Population by Continent by Year (2000)')+
  theme(axis.text.x = element_text(angle = 45))
print(pop_2000)
```

-   For the year 2000, the population for Africa exceeded that of Europe as it continues in its growth rate keeps rising compared to that of Europe, Asia still has the highest population.
